---
title: "Analysis"
output: pdf_document
---

```{r}
library(magrittr)
library(dplyr)
library(ggplot2)
```

```{r}
  teamsDf <- read.csv("professional-hockey-database/Teams.csv");
```

# Victory count

The first and easier analysis to do is check the team who won the most games.

```{r, fig.height=15}
  teamsDf %>%
  ggplot(aes(tmID, W)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip()
```

Plotting all the teams makes the graph too big and hard to read, so we might just get around 25% of the teams who won the most.

```{r}
  count <- nrow(teamsDf)
  subsetSize <- count*0.25
  mostWins <- teamsDf %>% arrange(-W) %>% tibble::rowid_to_column("rowNumber") %>% filter(rowNumber<subsetSize)
  
  mostWins %>% ggplot(aes(tmID, W)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

Having the teams with most wins is cool with our objective, but maybe it's a bit unfair. It seems to be a good idea to get the ratio between victories and losses in order to estimate the better team.

```{r}
  count <- nrow(teamsDf)
  subsetSize <- count*0.25
  bestWinRatio <- teamsDf %>% 
  mutate(WinRatio = W/G) %>%
  arrange(-WinRatio) %>% 
  tibble::rowid_to_column("rowNumber") %>% 
  filter(rowNumber<subsetSize)
  
  bestWinRatio %>% ggplot(aes(tmID, WinRatio)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

If we consider that the only negative result a team might have is a loss, we could also get the losses ratio. In this case, we are taking ties as results as good as victories, but it's still an interesting metric.

```{r}
  count <- nrow(teamsDf)
  subsetSize <- count*0.25
  bestLossRatio <- teamsDf %>% 
  mutate(LossRatio = L/G) %>%
  arrange(LossRatio) %>% 
  tibble::rowid_to_column("rowNumber") %>% 
  filter(rowNumber<subsetSize)
  
  bestLossRatio %>% ggplot(aes(tmID, LossRatio)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```